\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-04-25 forwards-and-backwards.tex
spellcheck

In what follows we first study various scaling limits of the spatial population process,
and then turn our attention to lineages traced back through these populations.
First, in Section~\ref{sec: Model and main results},
we describe the model and the main results,
Theorems~\ref{thm:nonlocal_convergence}, \ref{thm:local_convergence}, and~\ref{thm:lineages}.
Next, in Section~\ref{sec:applications}, we discuss a few striking consequences of these results
regarding the behavior of genealogies in traveling waves,
the appearance of periodic ``clumps'' in seemingly homogeneous population models,
and identifiability of the underlying dynamics from a stationary population profile.
In Section~\ref{sec:heuristics}, we provide heuristic explanations
of why the theorems ought to be true,
and some key ideas behind them,
and in Section~\ref{sec:lookdown} we define and discuss the lookdown construction.
Proofs of the results begin in Section~\ref{sec:proofs},
which proves results for population models with nonlocal interactions,
while Section~\ref{subsec:one step convergence proof} gives the more difficult proof
for the case when interaction distances also go to zero in the limit.
Finally, Section~\ref{sec:lookdown_proofs} gives proofs for convergence of the lookdown process
and the associated results for the motion of lineages.
The Appendix contains a few more techinical and less central lemmas.
The results are illustrated in a few places with individual-based simulations,
made using SLiM \citep{haller_slim_2019},
but these are provided for visualization and we do not embark on numerical study.

@@

Equation~(\ref{eqn:nonlocal_pde}) is a nonlocal version of a reaction-diffusion equation;
the diffusion is nonlinear if $\gamma$ depends on population density:
in other words, if the diffusivity of the population depends on the population density.
Passing to the classical limit, we recover equations like~(\ref{PME}).
Such equations are widely used in a number of contexts in biology in which
motility within a population varies with population density.
For example, density dependent dispersal is a common feature in spatial
models in ecology, eukaryotic cell biology, and avascular tumour growth;
see~\cite{sherratt:2010} and references therein for further discussion. 
In particular, it has been suggested as a model for
the expansion of a certain type of bacteria % of the type {\em Paenbacillus dendritiformis}
on a thin layer of agar in a petri dish 
\citep{cohen/golding/kozlovsky/benjacob/ron:1999}. 
We shall pay particular attention to the case in which the equation can be 
thought of as modelling the density of an expanding population. 
We focus instead on the monostable reaction of~(\ref{PME}).

@@

    Lines of descent can escape to infinite level in finite time, and so
    we endow $\lpmeasures$ with the vague topology ``in the level coordinate'',
    induced by test functions on $\overline{\IR^d} \times [0,\infty)$
    of the form $g(x) h(u)$, where $g \in C_b(\overline{\IR^d})$ is bounded and continuous
    and $h \in C_c([0, \infty))$ is compactly supported.
    This space is completely metrizable,
    and we may choose a countable set of nonnegative, compactly supported $f_k$
    such that a subset $K \subset \lpmeasures$ is relatively compact if and only if
    $\sup_{\lp \in K} \langle f_k, \lp \rangle < \infty$ for each $k$;
    below, Lemma~\ref{lem:compact containment lookdown} proves exatly this, and therefore compact containment.

    We require that $\{(F(\lp^N_t))_{t \ge 0}\}_N$
    is tight as a sequence of real-valued c\`adl\`ag processes, for all $F$ in a subset
    of $\bar C(\lpmeasures)$, the bounded, continuous functions on $\lpmeasures$,
    that is dense with respect to uniform convergence on compact subsets.
    Lemma~\ref{lem:test_fn_tightness} shows that $\{\langle f, \lp^N_t \rangle\}_N$ is a tight sequence
    for any $f: \overline{\IR^d} \times [0,\infty) \to \IR$ with compact support
    in the level direction.
    The functions $\lp \mapsto \langle f, \lp \rangle$
    do not form not a dense subset of $\bar C(\lpmeasures)$,
    but they do separate points and vanish nowhere, so
    by the Stone-Weierstrass theorem, the algebra they generate is dense
    % in $C_0(\lpmeasures)$, the continuous functions that vanish at infinity
    % (i.e., are less than $\epsilon$ outside of a compact set, for any $\epsilon$).
    % Since $C_0(\lpmeasures)$ is dense
    in $\bar C(\lpmeasures)$ (with respect to uniform convergence on compact subsets).
    So, it suffices to show that tightness for the processes $\langle f, \lp^N_t\rangle$
    extends to finite sums and products of these processes,
    which is shown in Lemma~\ref{lem:product_tightness}.
    The fact that martingale properties are preserved under passage to the limit
    is straightforward, and can be proved in a way analgous to Lemma~\ref{lem:limit_mgale};
    we omit the proof.
    Finally, we must show that the limiting lookdown process $\lp$
    projects to the limiting process $\eta$, i.e., a solution of the martingale
    problem in Theorem~\ref{thm:nonlocal_convergence}.
    Let $N_k \to \infty$ be a sequence along which $\lp^{N_k}$ converges.
    By Theorem~\ref{thm:nonlocal_convergence},
    there is a subsequence $N_{k(j)}$ along which the projected population processes
    $\eta^{N_{k(j)}}$ converge, and the limit solves the martingale problem.
    Thus any limit point of $\lp^N$ projects to a population process $\eta$
    solving the martingale problem of Theorem~\ref{thm:nonlocal_convergence}.
\end{proof}

@@

\begin{proof}[of Lemma~\ref{lem:h_containment}]
    Applied to $f(x, u) = h(u)$, the martinagle representation~\eqref{eqn:f_xi_mgale} is
    \begin{align*}
    \begin{split}
        \langle h, \lp_t^N \rangle
        &=
        \langle h, \lp_0^N \rangle
        + M^{h}_t 
        \\ & \qquad {}
        +
        \int_0^t
        \big\langle
            2 c_\theta(x, \eta^N_s) \int_u^N h(v) dv
        ,
        \lp_s^N(dx, du) \big \rangle
        ds
        \\ & \qquad \qquad {}
        +
        \int_0^t
        \big\langle
            \left(
                c_\theta(x, \eta^N_s) u^2
                - b_\theta(x, \eta^N_s) u
            \right)
            h'(u)
        ,
        \lp_s^N(dx, du) \big \rangle
        ds ,
    \end{split}
    \end{align*}
    where $M^{h}_t$ is a martingale
    with quadratic variation
    \begin{align*}
        [M^{h}]_t
        &=
        \int_0^t
        \langle 2 c_\theta(x, \eta_s^N) \int_u^N h(v)^2 dv
        ,
        \lp_s^N(dx, du) \big \rangle
        ds .
    \end{align*}
    Now, note that $0 \le c_\theta(x, \eta_x^N) \le C_a < \infty$
    and $b_\theta(x, \eta_s^N) \le C_b < \infty$,
    and we have assumed that $h'(u) \le 0$ (since $h$ is nonincreasing),
    so we may bound
    \begin{align} \label{eqn:hn_bound}
    \begin{split} 
        \langle h, \lp_t^N \rangle
        &\le
        \langle h, \lp_0^N \rangle
        + M^{h}_t
        \\ & \qquad {}
        +
        \int_0^t
        \big\langle
            2 C_a \int_u^\infty h(v) dv
            + \left( C_a u^2 + C_b u \right) |h'(u)|,
        \lp_s^N(dx, du) \big \rangle
        ds .
    \end{split}
    \end{align}
    Now, since $\lp_t^N$ is conditionally uniform given $\eta_t^N$,
    we know that for appropriate $f$,
    $  \IE[\langle f, \lp^N_t \rangle] = \IE[\langle \widetilde{f}, \eta^N_t \rangle] , $
    where $\widetilde{f}(x) = \int_0^\infty f(x, u) du$.
    By our assumptions on $h$,
    we know that
    $$
        \int_0^\infty \left( 2 C_a \int_u^\infty h(v) dv
        + \left( C_a u^2 + C_b u \right) |h'(u)| \right) du
        < C
    $$
    for some $C < \infty$, and so
    \begin{align*}
    \begin{split}
        \IE\left[ \langle h, \lp_t^N \rangle \right]
        &\le
        \IE\left[ \langle h, \lp_0^N \rangle \right]
        +
        C
        \int_0^t
        \IE\bigg[
            \big\langle
            1
            ,
            \eta_s^N \big \rangle
        \bigg]
        ds ,
    \end{split}
    \end{align*}
    which we know by Lemma~\ref{lem:eta_f_bound}
    is bounded by $C_0 e^{C_1 t}$ for some other constants $C_0$ and $C_1$.

@@

In this section we verify one of the conditions of the Markov Mapping Theorem,
namely, that ``integrating out levels'' in the generator of the lookdown process
we obtain the generator of the projected process.
In the notation of the thoerem,
we are verifying that $C$ defined in~\eqref{eqn:C_defn} is in fact $\Pgen^N$
(if defined with $A^N$) or $\Pgen^\infty$ (if defined with $A$).
We will work with test functions of the form
\begin{equation} \label{eqn:f_defn}
    f(\lp) = \prod_{(x, u) \in \lp} g(x, u) = \exp\left(\langle \log g, \lp \rangle \right) ,
\end{equation}
where $0 \le g \le 1$ and $g(x,u) = 1$ for all $u \ge u_g$ for some $u_g < \infty$.
Furthermore, recall that $\kappa^N(\lp)(\cdot) = \lp(\cdot \times [0, N)) / N$
is the ``spatial projection operator'',
and define the transition function $\Gamma^N : \measures \to \mathcal{M}(\IR^d \times [0,N))$
so that for $\eta \in \measures$, if $\hat g(x) = \int_0^N g(x, u) du / N$, then
\begin{align*}
    F^N_g(\eta)
    &:=
    \int f(\lp) \Gamma^N(\eta, d\lp) \\
    &=
    \exp\left(
        N \left\langle
            \log \frac{1}{N} \int_0^N g(x, u) du, \eta(dx)
        \right\rangle
    \right) \\
    &=
    \exp\left(
        N \left\langle \log \hat g(x), \eta(dx) \right\rangle
    \right) ,
\end{align*}
i.e., $\Gamma^N$ assigns independent labels on $[0, N]$ to each of the points in $\eta$.
It follows from Lemma~\ref{def: MP definition of limit} that for test functions of this form
the generator of $\eta^N_t$ is
\begin{align} \label{eqn:pgen_defn_again}
    \begin{split}
    \Pgen^N F^N_g(\eta)
    =
    F^N_g(\eta)
        N \theta & \bigg\langle
        \gamma(x, \eta) \int r(z, \eta) \left( \hat g(z) - 1 \right) q_\theta(x, dz)
    \\ &\qquad \qquad {}
        +
        \mu_\theta(x, \eta) \left( \frac{1}{\hat g(x)} - 1 \right)
        ,
        \eta(dx)
    \bigg\rangle .
    \end{split}
\end{align}
(Note that $f$ here differs from the $f$ used in Lemma~\ref{def: MP definition of limit}
so as to agree with standard usage in the literature on lookdown processes.)
The generator of $\lp^N_t$ is $A^N$, defined in equation~\eqref{eqn:lookdown_generator}.

@@

    Next, note that for $0 \le \tau_n \le T$,
    \begin{align*}
        &
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta} Y^{(n)}_{\tau_n + \theta} 
            -
            X^{(n)}_{\tau_n} Y^{(n)}_{\tau_n} 
        \right|
        \\ &\qquad 
        \le
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left| X^{(n)}_{\tau_n + \theta} \right|
        \left|
            Y^{(n)}_{\tau_n + \theta} 
            -
            Y^{(n)}_{\tau_n} 
        \right|
        +
        \left|
            X^{(n)}_{\tau_n + \theta}
            -
            X^{(n)}_{\tau_n}
        \right|
        \left| Y^{(n)}_{\tau_n} \right|
        \\ &\qquad 
        \le
        \sup_{0 \le t \le T}
            \left| X^{(n)}_t \right|
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            Y^{(n)}_{\tau_n + \theta} 
            -
            Y^{(n)}_{\tau_n} 
        \right|
        +
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta}
            -
            X^{(n)}_{\tau_n}
        \right|
        \sup_{0 \le t \le T}
        \left| Y^{(n)}_t \right| ,
    \end{align*}
    so that for any $C$,
    \begin{align} \label{eqn:xy_bound}
        \begin{split}
        &
        \IP\left\{
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta} Y^{(n)}_{\tau_n + \theta} 
            -
            X^{(n)}_{\tau_n} Y^{(n)}_{\tau_n} 
        \right|
        > \epsilon
        \right\}
    \\ &\qquad \le
        \IP\left\{
            \sup_{0 \le t \le T}
                \left| X^{(n)}_t \right|
            > C
        \right\}
    +
        \IP\left\{
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            Y^{(n)}_{\tau_n + \theta} 
            -
            Y^{(n)}_{\tau_n} 
        \right|
            > \epsilon/C
        \right\}
    \\ &\qquad {} +
        \IP\left\{
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta} 
            -
            X^{(n)}_{\tau_n} 
        \right|
            > \epsilon/C
        \right\}
    +
        \IP\left\{
            \sup_{0 \le t \le T}
                \left| Y^{(n)}_t \right|
            > C
        \right\}
        \end{split}
    \end{align}
    Now, since $\max_{0 \le t \le T} X^{(n)}_t$ is tight (and likewise for $Y$)
    (see, e.g., Remark 3.7.3 in \citet{ethier/kurtz:1986}),
    we may choose a $C \ge 4$ for which
    $$
    \IP\left\{
        \sup_{0 \le t \le T}
            \left| X^{(n)}_t \right|
        > C
    \right\} \le \frac{\epsilon}{4} .
    $$
    Simliarly, by assumption we can choose a $\delta$ for which
    $$
        \IP\left\{
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta} 
            -
            X^{(n)}_{\tau_n} 
        \right|
            > \epsilon/C
        \right\}
        \le \frac{\epsilon}{C} .
    $$
    If we choose $C$ and $\delta$ that do this for both $X^{(n)}$ and $Y^{(n)}$,
    then each of the terms in equation~\eqref{eqn:xy_bound} are bounded by $\epsilon/4$,
    and condition (2) is satisfied for the product process.
\end{proof}

>>>>

In what follows we first study various scaling limits of the spatial population process,
and then turn our attention to lineages traced back through these populations.
First, in Section~\ref{sec: Model and main results},
we describe the model and the main results,
Theorems~\ref{thm:nonlocal_convergence}, \ref{thm:local_convergence}, and~\ref{thm:lineages}.
Next, in Section~\ref{sec:applications}, we discuss a few striking consequences of these results
regarding the behavior of genealogies in traveling waves,
the appearance of periodic ``clumps'' in seemingly homogeneous population models,
and identifiability of the underlying dynamics from a stationary population profile.
In Section~\ref{sec:heuristics}, we provide heuristic explanations
of why the theorems ought to be true,
and some key ideas behind them,
and in Section~\ref{sec:lookdown} we define and discuss the lookdown construction.
Proofs of the results begin in Section~\ref{sec:proofs},
which proves results for population models with nonlocal interactions,
while Section~\ref{subsec:one step convergence proof} gives the more difficult proof
for the case when interaction distances also go to zero in the limit.
Finally, Section~\ref{sec:lookdown_proofs} gives proofs for convergence of the lookdown process
and the associated results for the motion of lineages.
The Appendix contains a few more technical and less central lemmas.
The results are illustrated in a few places with individual-based simulations,
made using SLiM \citep{haller_slim_2019},
but these are provided for visualization and we do not embark on numerical study.

@@

Equation~(\ref{eqn:nonlocal_pde}) is a nonlocal version of a reaction-diffusion equation;
the diffusion is nonlinear if $\gamma$ depends on population density:
in other words, if the diffusivity of the population depends on the population density.
Passing to the classical limit, we recover equations like~(\ref{PME}).
Such equations are widely used in a number of contexts in biology in which
motility within a population varies with population density.
For example, density dependent dispersal is a common feature in spatial
models in ecology, eukaryotic cell biology, and avascular tumour growth;
see~\cite{sherratt:2010} and references therein for further discussion. 
In particular, it has been suggested as a model for
the expansion of a certain type of bacteria % of the type {\em Paenbacillus dendritiformis}
on a thin layer of agar in a Petri dish 
\citep{cohen/golding/kozlovsky/benjacob/ron:1999}. 
We shall pay particular attention to the case in which the equation can be 
thought of as modelling the density of an expanding population. 
We focus instead on the monostable reaction of~(\ref{PME}).

@@

    Lines of descent can escape to infinite level in finite time, and so
    we endow $\lpmeasures$ with the vague topology ``in the level coordinate'',
    induced by test functions on $\overline{\IR^d} \times [0,\infty)$
    of the form $g(x) h(u)$, where $g \in C_b(\overline{\IR^d})$ is bounded and continuous
    and $h \in C_c([0, \infty))$ is compactly supported.
    This space is completely metrizable,
    and we may choose a countable set of nonnegative, compactly supported $f_k$
    such that a subset $K \subset \lpmeasures$ is relatively compact if and only if
    $\sup_{\lp \in K} \langle f_k, \lp \rangle < \infty$ for each $k$;
    below, Lemma~\ref{lem:compact containment lookdown} proves exactly this, and therefore compact containment.

    We require that $\{(F(\lp^N_t))_{t \ge 0}\}_N$
    is tight as a sequence of real-valued c\`adl\`ag processes, for all $F$ in a subset
    of $\bar C(\lpmeasures)$, the bounded, continuous functions on $\lpmeasures$,
    that is dense with respect to uniform convergence on compact subsets.
    Lemma~\ref{lem:test_fn_tightness} shows that $\{\langle f, \lp^N_t \rangle\}_N$ is a tight sequence
    for any $f: \overline{\IR^d} \times [0,\infty) \to \IR$ with compact support
    in the level direction.
    The functions $\lp \mapsto \langle f, \lp \rangle$
    do not form not a dense subset of $\bar C(\lpmeasures)$,
    but they do separate points and vanish nowhere, so
    by the Stone-Weierstrass theorem, the algebra they generate is dense
    % in $C_0(\lpmeasures)$, the continuous functions that vanish at infinity
    % (i.e., are less than $\epsilon$ outside of a compact set, for any $\epsilon$).
    % Since $C_0(\lpmeasures)$ is dense
    in $\bar C(\lpmeasures)$ (with respect to uniform convergence on compact subsets).
    So, it suffices to show that tightness for the processes $\langle f, \lp^N_t\rangle$
    extends to finite sums and products of these processes,
    which is shown in Lemma~\ref{lem:product_tightness}.
    The fact that martingale properties are preserved under passage to the limit
    is straightforward, and can be proved in a way analogous to Lemma~\ref{lem:limit_mgale};
    we omit the proof.
    Finally, we must show that the limiting lookdown process $\lp$
    projects to the limiting process $\eta$, i.e., a solution of the martingale
    problem in Theorem~\ref{thm:nonlocal_convergence}.
    Let $N_k \to \infty$ be a sequence along which $\lp^{N_k}$ converges.
    By Theorem~\ref{thm:nonlocal_convergence},
    there is a subsequence $N_{k(j)}$ along which the projected population processes
    $\eta^{N_{k(j)}}$ converge, and the limit solves the martingale problem.
    Thus any limit point of $\lp^N$ projects to a population process $\eta$
    solving the martingale problem of Theorem~\ref{thm:nonlocal_convergence}.
\end{proof}

@@

\begin{proof}[of Lemma~\ref{lem:h_containment}]
    Applied to $f(x, u) = h(u)$, the martingale representation~\eqref{eqn:f_xi_mgale} is
    \begin{align*}
    \begin{split}
        \langle h, \lp_t^N \rangle
        &=
        \langle h, \lp_0^N \rangle
        + M^{h}_t 
        \\ & \qquad {}
        +
        \int_0^t
        \big\langle
            2 c_\theta(x, \eta^N_s) \int_u^N h(v) dv
        ,
        \lp_s^N(dx, du) \big \rangle
        ds
        \\ & \qquad \qquad {}
        +
        \int_0^t
        \big\langle
            \left(
                c_\theta(x, \eta^N_s) u^2
                - b_\theta(x, \eta^N_s) u
            \right)
            h'(u)
        ,
        \lp_s^N(dx, du) \big \rangle
        ds ,
    \end{split}
    \end{align*}
    where $M^{h}_t$ is a martingale
    with quadratic variation
    \begin{align*}
        [M^{h}]_t
        &=
        \int_0^t
        \langle 2 c_\theta(x, \eta_s^N) \int_u^N h(v)^2 dv
        ,
        \lp_s^N(dx, du) \big \rangle
        ds .
    \end{align*}
    Now, note that $0 \le c_\theta(x, \eta_x^N) \le C_a < \infty$
    and $b_\theta(x, \eta_s^N) \le C_b < \infty$,
    and we have assumed that $h'(u) \le 0$ (since $h$ is nonincreasing),
    so we may bound
    \begin{align} \label{eqn:hn_bound}
    \begin{split} 
        \langle h, \lp_t^N \rangle
        &\le
        \langle h, \lp_0^N \rangle
        + M^{h}_t
        \\ & \qquad {}
        +
        \int_0^t
        \big\langle
            2 C_a \int_u^\infty h(v) dv
            + \left( C_a u^2 + C_b u \right) |h'(u)|,
        \lp_s^N(dx, du) \big \rangle
        ds .
    \end{split}
    \end{align}
    Now, since $\lp_t^N$ is conditionally uniform given $\eta_t^N$,
    we know that for appropriate $f$,
    $  \IE[\langle f, \lp^N_t \rangle] = \IE[\langle \widetilde{f}, \eta^N_t \rangle] , $
    where $\widetilde{f}(x) = \int_0^\infty f(x, u) du$.
    By our assumptions on $h$,
    we know that
    $$
        \int_0^\infty \left( 2 C_a \int_u^\infty h(v) dv
        + \left( C_a u^2 + C_b u \right) |h'(u)| \right) du
        < C
    $$
    for some $C < \infty$, and so
    \begin{align*}
    \begin{split}
        \IE\left[ \langle h, \lp_t^N \rangle \right]
        &\le
        \IE\left[ \langle h, \lp_0^N \rangle \right]
        +
        C
        \int_0^t
        \IE\bigg[
            \big\langle
            1
            ,
            \eta_s^N \big \rangle
        \bigg]
        ds ,
    \end{split}
    \end{align*}
    which we know by Lemma~\ref{lem:eta_f_bound}
    is bounded by $C_0 e^{C_1 t}$ for some other constants $C_0$ and $C_1$.

@@

In this section we verify one of the conditions of the Markov Mapping Theorem,
namely, that ``integrating out levels'' in the generator of the lookdown process
we obtain the generator of the projected process.
In the notation of the theorem,
we are verifying that $C$ defined in~\eqref{eqn:C_defn} is in fact $\Pgen^N$
(if defined with $A^N$) or $\Pgen^\infty$ (if defined with $A$).
We will work with test functions of the form
\begin{equation} \label{eqn:f_defn}
    f(\lp) = \prod_{(x, u) \in \lp} g(x, u) = \exp\left(\langle \log g, \lp \rangle \right) ,
\end{equation}
where $0 \le g \le 1$ and $g(x,u) = 1$ for all $u \ge u_g$ for some $u_g < \infty$.
Furthermore, recall that $\kappa^N(\lp)(\cdot) = \lp(\cdot \times [0, N)) / N$
is the ``spatial projection operator'',
and define the transition function $\Gamma^N : \measures \to \mathcal{M}(\IR^d \times [0,N))$
so that for $\eta \in \measures$, if $\hat g(x) = \int_0^N g(x, u) du / N$, then
\begin{align*}
    F^N_g(\eta)
    &:=
    \int f(\lp) \Gamma^N(\eta, d\lp) \\
    &=
    \exp\left(
        N \left\langle
            \log \frac{1}{N} \int_0^N g(x, u) du, \eta(dx)
        \right\rangle
    \right) \\
    &=
    \exp\left(
        N \left\langle \log \hat g(x), \eta(dx) \right\rangle
    \right) ,
\end{align*}
i.e., $\Gamma^N$ assigns independent labels on $[0, N]$ to each of the points in $\eta$.
It follows from Lemma~\ref{def: MP definition of limit} that for test functions of this form
the generator of $\eta^N_t$ is
\begin{align} \label{eqn:pgen_defn_again}
    \begin{split}
    \Pgen^N F^N_g(\eta)
    =
    F^N_g(\eta)
        N \theta & \bigg\langle
        \gamma(x, \eta) \int r(z, \eta) \left( \hat g(z) - 1 \right) q_\theta(x, dz)
    \\ &\qquad \qquad {}
        +
        \mu_\theta(x, \eta) \left( \frac{1}{\hat g(x)} - 1 \right)
        ,
        \eta(dx)
    \bigg\rangle .
    \end{split}
\end{align}
(Note that $f$ here differs from the $f$ used in Lemma~\ref{def: MP definition of limit}
so as to agree with standard usage in the literature on lookdown processes.)
The generator of $\lp^N_t$ is $A^N$, defined in equation~\eqref{eqn:lookdown_generator}.

@@

    Next, note that for $0 \le \tau_n \le T$,
    \begin{align*}
        &
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta} Y^{(n)}_{\tau_n + \theta} 
            -
            X^{(n)}_{\tau_n} Y^{(n)}_{\tau_n} 
        \right|
        \\ &\qquad 
        \le
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left| X^{(n)}_{\tau_n + \theta} \right|
        \left|
            Y^{(n)}_{\tau_n + \theta} 
            -
            Y^{(n)}_{\tau_n} 
        \right|
        +
        \left|
            X^{(n)}_{\tau_n + \theta}
            -
            X^{(n)}_{\tau_n}
        \right|
        \left| Y^{(n)}_{\tau_n} \right|
        \\ &\qquad 
        \le
        \sup_{0 \le t \le T}
            \left| X^{(n)}_t \right|
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            Y^{(n)}_{\tau_n + \theta} 
            -
            Y^{(n)}_{\tau_n} 
        \right|
        +
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta}
            -
            X^{(n)}_{\tau_n}
        \right|
        \sup_{0 \le t \le T}
        \left| Y^{(n)}_t \right| ,
    \end{align*}
    so that for any $C$,
    \begin{align} \label{eqn:xy_bound}
        \begin{split}
        &
        \IP\left\{
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta} Y^{(n)}_{\tau_n + \theta} 
            -
            X^{(n)}_{\tau_n} Y^{(n)}_{\tau_n} 
        \right|
        > \epsilon
        \right\}
    \\ &\qquad \le
        \IP\left\{
            \sup_{0 \le t \le T}
                \left| X^{(n)}_t \right|
            > C
        \right\}
    +
        \IP\left\{
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            Y^{(n)}_{\tau_n + \theta} 
            -
            Y^{(n)}_{\tau_n} 
        \right|
            > \epsilon/C
        \right\}
    \\ &\qquad {} +
        \IP\left\{
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta} 
            -
            X^{(n)}_{\tau_n} 
        \right|
            > \epsilon/C
        \right\}
    +
        \IP\left\{
            \sup_{0 \le t \le T}
                \left| Y^{(n)}_t \right|
            > C
        \right\}
        \end{split}
    \end{align}
    Now, since $\max_{0 \le t \le T} X^{(n)}_t$ is tight (and likewise for $Y$)
    (see, e.g., Remark 3.7.3 in \citet{ethier/kurtz:1986}),
    we may choose a $C \ge 4$ for which
    $$
    \IP\left\{
        \sup_{0 \le t \le T}
            \left| X^{(n)}_t \right|
        > C
    \right\} \le \frac{\epsilon}{4} .
    $$
    Similarly, by assumption we can choose a $\delta$ for which
    $$
        \IP\left\{
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta} 
            -
            X^{(n)}_{\tau_n} 
        \right|
            > \epsilon/C
        \right\}
        \le \frac{\epsilon}{C} .
    $$
    If we choose $C$ and $\delta$ that do this for both $X^{(n)}$ and $Y^{(n)}$,
    then each of the terms in equation~\eqref{eqn:xy_bound} are bounded by $\epsilon/4$,
    and condition (2) is satisfied for the product process.
\end{proof}
